COURSERA-Reproducible Research Course Project 1
===============================================

Name: Christoph Thommen
Date: 17.10.2015
Data: The dataset contains information about the number of steps taken by 
an individual collected during October and November 2012 in a 5 minute interval.
The following variables are collected:
steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##Loading and preprocessing the data

```{r}
raw_data = read.csv("activity.csv")
```

```{r}
raw_data$date=as.date(raw_data$date)
```
#stimmt nicht -> checken

##What is mean total number of steps taken per day?

```{r}
result1=aggregate(raw_data$steps,by=list(raw_data$date),sum)
dimnames(result1)[[2]]=c("date","steps")
```

```{r}
hist(result1$steps)
```

```{r}
mean(result1$steps,na.rm=T)
median(result1$steps,na.rm=T)
```
#What is the average daily activity pattern?

```{r}
plot(result1$steps,type="l")
```

```{r}
result1[max(result1$steps,na.rm=T),]$date
```
#checken -> stimmt nicht

##Imputing missing values

```{r}
table(is.na(raw_data$steps)==T)
table(is.na(raw_data==T))
```

```{r}
result2=aggregate(raw_data$steps,by=list(raw_data$interval),mean,na.rm=T)
result2$steps=round(result2$steps,0)
dimnames(result2)[[2]]=c("interval","steps")
```

raw_data_imp=raw_data[is.na(raw_data$steps)==T,]
#raw_data_imp[raw_data_imp$interval==result2$interval,]$steps=result2[raw_data_imp$interval==result2$interval,]$steps

raw_data_imp2=merge(raw_data_imp,result2,by="interval",all.x=T)

raw_data_imp3=raw_data_imp2[,c(1,3,4)]
dimnames(raw_data_imp3)[[2]]=c("interval","date","steps")

raw_data2=rbind(raw_data[is.na(raw_data$steps)==F,],raw_data_imp3)


#raw_data$steps=ifelse(is.na(raw_data$steps)==T,result2[raw_data$interval==result2$interval,]$steps,raw_data$steps)
hist(raw_data2$steps)

mean(raw_data2$steps,na.rm=T)
median(raw_data2$steps,na.rm=T)


##Are there differences in activity patterns between weekdays and weekends?
#date als Datum definieren
raw_data2$weekday=ifelse(raw_data2$weekdays %in% c("Sunday","Saturday"),"weekend","weekday")


plot(result1$steps,type="l")


knit2html("PA1.Rmd")
